Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India: vertical stratification reflects guild-dependent dynamics

### 2016-08-09 inst/doc/draft.Rmd
v0.1.9031 Strata analysis

---
title: 'Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India: vertical stratification reflects guild-dependent dynamics'
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output: word_document
csl: mee.csl
bibliography: refs.bib
---

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Trees <- genTrees()
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp)
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
XY <- Trees[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
Trees$com <- (XY %over% com)[,1]
com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
names(Deriv)[5] <- 'SW'
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(R) <- com@data$id
L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
L[is.na(L)] <- 0
Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')]
row.names(Q) <- PFT_sp$SpCode
sel <- intersect(row.names(Q), names(L))
Q <- Q[sel,]
L <- L[sel]
R <- R[row.names(L),]
L <- list(abundance = L, `presence-absence` = L)
L$`presence-absence`[L$`presence-absence` > 1]  <- 1
rm(Deriv, XY, com, Env, Trees, sel, BA)
```

@@

```{r CWM}
# Observed CWMs
CWM <- lapply(L, function(l){lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){apply(l, 1, function(x){weighted.mean(Q[,v], x, na.rm=T)})})})
CWM <- lapply(CWM, function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT') ; return(x)})
## Regressions
reg <- paste('~', paste(names(R), collapse = ' + '))
CWM.lm <- lapply(CWM, function(x){lapply(x, function(y){lm(y ~., R)})})
CWM.lm <- lapply(CWM.lm, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
## Global F
Fval <- lapply(CWM.lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})
Fval <- lapply(Fval, function(x){do.call('rbind.data.frame', x)})
Fval <- data.frame(Fval)
names(Fval) <- names(CWM.lm)
## Partial F
pFval <- lapply(CWM.lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})

# Simulated CWMs
n = 999
nQ <- rep(list(Q), n)
nQ <- lapply(nQ, function(x){row.names(x) <- sample(row.names(x)) ; return(x)})
nQ <- lapply(nQ, function(x){x[names(L$abundance),]})
cl <- makeCluster(cores)
clusterExport(cl, list('L', 'nQ'))
nCWM <- clusterApply(cl, L, function(l){
  lapply(nQ, function(q){
    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
      apply(l, 1, function(x){
        weighted.mean(q[,v], x, na.rm=T)
      })
    })
  })
})
stopCluster(cl)
rm(cl)
nCWM <- lapply(nCWM, function(y){lapply(y,function(x){names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT') ; return(x)})})
## Regressions
nCWM.lm <- mapply(function(x,y){
  lapply(y, function(z){
    mapply(function(r,s){
      lm(formula(paste0('s ~ ', as.character(r$terms)[3])), cbind(s,R))
    }, r = x, s = z, SIMPLIFY = F)
  })
}, x = CWM.lm, y = nCWM, SIMPLIFY = F)
rm(nCWM) ; invisible(gc())
## Global F
nFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){summary(x)$fstatistic['value']})})})
nFval <- lapply(nFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
nFval <- lapply(nFval, function(f){lapply(f, unlist)})
pval <- 1 - mapply(function(x,y){mapply(function(r,s){rank(c(r,s))[1]/(n+1)}, r = x, s= y)}, x = lapply(as.list(Fval), as.list), y = nFval)
row.names(pval) <- row.names(Fval)
rm(Fval, nFval)
# Partial F
npFval <- lapply(nCWM.lm, function(lm){lapply(lm, function(y){lapply(y, function(x){car::Anova(x)['F value']})})})
npFval <- lapply(npFval, function(x){lapply(x, `[`, names(x[[1]])) ; apply(do.call(rbind, x), 2, as.list)})
npFval <- lapply(npFval, function(f){lapply(f, function(x){do.call('cbind', x)})})
ppval <- mapply(function(x,y){mapply(function(r,s){mapply(function(p,q){rank(c(p,q))[1]/(n+1)}, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))}, r = x, s = y)}, x = pFval, y = npFval, SIMPLIFY = F)
ppval <- lapply(ppval, function(y){lapply(y, function(x){data.frame(t(x))})})
ppval <- lapply(ppval, function(y){data.frame(data.table::rbindlist(y, fill = T))})
ppval <- lapply(ppval, function(y){row.names(y) <- names(CWM.lm$abundance) ; return(y)})
ppval$abundance <- ppval$abundance[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy')]
ppval$abundance$BA <- NA
ppval$`presence-absence` <- ppval$`presence-absence`[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
ppval <- lapply(ppval, function(x){1 - x})
rm(pFval, npFval)

# Result tables
t1 <- regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]])
t1 <- cbind(c('AB', rep(' ', length(t1[,1])-1)), t1)
names(t1)[1] <- ' '
t2 <- regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]])
t2 <- cbind(c('PA', rep(' ', length(t2[,2])-1)), t2)
names(t2)[1] <- ' '
kable(bold(rbind(t1,t2)), row.names = F)
rm(CWM.lm, nCWM.lm, pval, ppval, reg, t1, t2) ; invisible(gc())
```

Traits convergence/divergence was measured with community weighted variances (CWV). Overall, observed CWVs DID NOT differ from null expectations for all traits. We could found significant relationships between standardized effect size (SES) of CWVs and environmental variables (see table 4). Convergence of LES and LDMC and divergence of LA were found under south-west orientation. LA and LDMC values also diverged in communities with increasing curvature. Observed CVNNDs did not reveal niche differentiation among communities for all traits. But we also found significant relationships of their standardized effect size (SES) of CVNNDs with environmental variables (see table 4). niche differentiation of LA increased in lower curvature and wetness contexts, while differentiation of WD decreased in southWest orientation.

@@

```{r CWV}
# Observed CWVs
CWV <- lapply(names(CWM$abundance), function(v){apply(cbind(data.frame(CWM$abundance[[which(names(CWM$abundance) == v)]]),L$abundance), 1, function(x){sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))})})
names(CWV) <- names(CWM$abundance)

# Simulated CWVs 
nL <- rep(list(L$abundance), n)
nL <- lapply(nL,function(l){apply(l, 1, function(x){x[x != 0] <- sample(x[x != 0]) ; return(x)})}) 
nL <- lapply(nL, function(x){t(x)})
nL <- lapply(nL, data.frame)

cl <- makeCluster(cores)
clusterExport(cl, list('nL', 'CWM', 'Q'))
nCWM.cwv <- clusterApply(cl, nL, function(l){
  lapply(as.list(names(CWM$abundance)), function(v){
    apply(l, 1, function(x){
      weighted.mean(Q[,v], x, na.rm=T)
    })
  })
})
stopCluster(cl)
nCWM.cwv <- lapply(nCWM.cwv, function(x){names(x) <- names(CWM$abundance) ; return(x)})
nCWV <- mapply(function(l,cwm){
  lapply(names(CWM$abundance), function(v){
    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
      sqrt(weighted.mean((Q[,v]-x[1])^2, x[-1], na.rm=T))
    })
  })
}, l = nL, cwm = nCWM.cwv, SIMPLIFY = F)
nCWV <- lapply(nCWV, function(x){names(x) <- names(CWM$abundance) ; return(x)})
rm(nCWM.cwv, cl, nL) ; invisible(gc())
## Non null CWVs
nCWV <- lapply(nCWV, `[`, names(nCWV[[1]]))
nCWV <- apply(do.call(rbind, nCWV), 2, as.list)
nCWV <- lapply(nCWV, function(x){do.call('cbind', x)})
nCWV <- lapply(nCWV, data.frame)
CWV.pval <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    rank(c(p,q))[1]/(n+1)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.pval.res <- data.frame(inf = apply(apply(CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}), sup = apply(apply(1 - CWV.pval, 2, p.adjust, method = 'fdr'), 2, function(x){length(which(x < 0.025))/length(x)}))
## CWV SES
CWV.ses <- data.frame(mapply(function(x,y){
  mapply(function(p,q){
    (p - mean(q))/sd(q)
  }, p = as.list(x), q = as.list(data.frame(t(y))))
}, x = CWV, y = nCWV))
CWV.ses.lm <- apply(CWV.ses, 2, function(y){lm(y ~., R)})
CWV.ses.lm <- lapply(CWV.ses.lm, function(x){step(x, direction = 'both', trace = 0)})
CWV.ses.res <- lapply(CWV.ses.lm, function(x){summary(x)$coefficients})
CWV.ses.res <- lapply(CWV.ses.res, data.frame)
CWV.ses.res <- lapply(CWV.ses.res, function(x){x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   ')) ; return(x)})
CWV.ses.res <- data.frame(data.table::rbindlist(lapply(CWV.ses.res, function(x){data.frame(t(x[5]))}), fill = T))
CWV.ses.res <- CWV.ses.res[-1]
var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
CWV.ses.res[,setdiff(var, names(CWV.ses.res))] <- NA
CWV.ses.res <- CWV.ses.res[var]
CWV.ses.res <- apply(CWV.ses.res, 2, as.character)
row.names(CWV.ses.res) <- names(CWV.ses)
CWV.ses.res[is.na(CWV.ses.res)] <- ''
CWV.ses.res <- data.frame(CWV.ses.res)
CWV.ses.res$`$R^2$` <- round(unlist(lapply(CWV.ses.lm, function(x){summary(x)$r.squared})),3)

# Results
# barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
CWV.ses.res <- cbind(c('CWV', rep(' ', length(CWV.ses.res[,1])-1)), row.names(CWV.ses.res), CWV.ses.res)
names(CWV.ses.res)[1:2] <- c(' ', 'Trait')
rm(CWV.pval, CWV.pval.res, CWV.ses.lm, nCWV)
```

```{r CVNND}
# Observed CVNND
CVNND <- apply(L$abundance, 1, function(l){apply(Q, 2, function(q){cati::CVNND(rep(q, l), na.rm = T)})})

>>>>

---
title: 'Fine scale determinants of wood and leaf strategies in a low-elevation rainforest of India: vertical stratification reflects guild-dependent dynamics'
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output: pdf_document
csl: mee.csl
bibliography: refs.bib
---

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Trees <- genTrees()
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp)
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
XY <- Trees[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
Trees$com <- (XY %over% com)[,1]
com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'cosaspect'))
names(Deriv)[5] <- 'SW'
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
BA <- aggregate(Trees$Girth, list(Trees$com), function(x){sum(pi*(x/(2*pi))^2, na.rm = T)})
com@data$BA <- BA$x[match(com@data$id, BA$Group.1)]
pca.env <- princomp(~ Elevation + Slope + Curvature + PlanCurvature + ProfileCurvature + Wetness + SW, data = com@data, cor = T)
R <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(R) <- com@data$id
L <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
L[is.na(L)] <- 0
Q <- PFT_sp[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')]
row.names(Q) <- PFT_sp$SpCode
sel <- which(Species$Strata[match(PFT_sp$SpCode, Species$SpCode)] %in% c('ESI', 'ESII'))
Q <- list(C = Q[sel,], U = Q[-sel,])
sel <- lapply(Q, function(q){intersect(row.names(q), names(L))})
Q <- mapply(function(q,s){q[s,]}, q = Q, s = sel, SIMPLIFY = F)
L <- lapply(sel, function(s){L[s]})
R <- lapply(L, function(l){R[row.names(l),]})
L <- lapply(L, function(l){list(abundance = l, `presence-absence` = l)})
L <- lapply(L, function(l){l$`presence-absence`[l$`presence-absence` > 1]  <- 1 ; return(l)})
rm(Deriv, XY, Env, sel, BA)
```

@@

```{r CWM}
# Observed CWMs
CWM <- mapply(function(Ls,Qs){
  lapply(Ls, function(l){
    lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
      apply(l, 1, function(x){
        weighted.mean(Qs[,v], x, na.rm=T)
      })
    })
  })
}, Ls = L, Qs = Q, SIMPLIFY = F)
CWM <- lapply(CWM, function(s){
  lapply(s, function(x){
    names(x) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')
    return(x)
  })
})
## Regressions
CWM.lm <- mapply(function(CWMs, Rs){
  lapply(CWMs, function(p){
    lapply(p, function(t){
      lm(t ~., Rs)
    })
  })
}, CWMs = CWM, Rs = R, SIMPLIFY = F)
CWM.lm <- mapply(function(s, Rs){
  lapply(s, function(p){
    lapply(p, function(t){
      step(t, direction = 'both', trace = 0)
    })
  })
}, s = CWM.lm, Rs = R, SIMPLIFY = F)
## Global F
Fval <- lapply(CWM.lm, function(s){
  lapply(s, function(y){
    lapply(y, function(x){
      summary(x)$fstatistic['value']
    })
  })
})
Fval <- lapply(Fval, function(s){
  lapply(s, function(x){
    do.call('rbind.data.frame', x)
  })
})
Fval <- lapply(Fval, function(s){
  s <- data.frame(s)
  names(s) <- names(CWM.lm$C)
  return(s)
})
Fval$U <- Fval$U[c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),]
## Partial F
pFval <- lapply(CWM.lm, function(s){
  lapply(s, function(y){
    lapply(y, function(x){
      car::Anova(x)['F value']
    })
  })
})

# Simulated CWMs
n = 999
nQ <- lapply(Q, function(s){
  rep(list(s), n)
})
nQ <- lapply(nQ, function(s){ 
  lapply(s, function(nq){
      row.names(nq) <- sample(row.names(nq))
      return(nq)
  })
})
nQ <- mapply(function(nQs, Ls){
  lapply(nQs, function(nq){
    nq[names(Ls$abundance),]
  })
}, nQs = nQ, Ls = L, SIMPLIFY = F)
cl <- makeCluster(cores)
clusterExport(cl, list('L', 'nQ'))
nCWM <- clusterMap(cl, function(nQs, Ls){
  lapply(nQs, function(nq){
    lapply(Ls, function(l){
      lapply(as.list(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')), function(v){
        apply(l, 1, function(x){
          weighted.mean(nq[,v], x, na.rm=T)
        })
      })
    })
  })
}, Ls = L, nQs = nQ)
stopCluster(cl)
rm(cl)
nCWM <- lapply(nCWM, function(s){
  lapply(s, function(ns){
  lapply(ns,function(ab){
    names(ab) <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT')
    return(ab)
    })
  })
})
## Regressions
nCWM.lm <- mapply(function(CWM.lms, nCWMs, Rs){
  lapply(nCWMs, function(ncwm){
    mapply(function(CWM.lmab, ncwmab){
      mapply(function(CWM.lmt, ncwmt){
        lm(formula(paste0('ncwmt ~ ', as.character(CWM.lmt$terms)[3])), cbind(ncwmt,Rs))
      }, CWM.lmt = CWM.lmab, ncwmt = ncwmab, SIMPLIFY = F)
    }, CWM.lmab = CWM.lms, ncwmab = ncwm, SIMPLIFY = F)
  })
}, CWM.lms = CWM.lm, nCWMs = nCWM, Rs = R, SIMPLIFY = F)
rm(nCWM)
invisible(gc())
## Global F
nFval <- lapply(nCWM.lm, function(nCWM.lms){
  lapply(nCWM.lms, function(lm){
    lapply(lm, function(y){
      lapply(y, function(x){
        summary(x)$fstatistic['value']
      })
    })
  })
})
nFval <- lapply(nFval, function(x){
  lapply(x, `[`, names(x[[1]]))
  apply(do.call(rbind, x), 2, as.list)
})
nFval <- lapply(nFval, function(s){
  lapply(s, function(x){
    lapply(x, `[`, names(x[[1]]))
    apply(do.call(rbind, x), 2, as.list)
  })
})
nFval <- lapply(nFval, function(s){
  lapply(s, function(ab){
    lapply(ab, unlist)
  })
})
pval <- mapply(function(Fvals, nFvals){
  1 -  mapply(function(Fvalab, nFvalab){
    mapply(function(Fvalt,nFvalt){
      if(!is.na(Fvalt))
        rank(c(Fvalt, nFvalt))[1]/(n+1)
      else
        NA
    }, Fvalt = Fvalab, nFvalt = nFvalab)
  }, Fvalab = lapply(as.list(Fvals), as.list), nFvalab = nFvals)
}, Fvals = Fval, nFvals = nFval, SIMPLIFY = F)
pval <- lapply(pval, function(s){
  row.names(s) <- row.names(Fval$C)
  return(s)
})
rm(Fval, nFval)
# Partial F
npFval <- lapply(nCWM.lm, function(s){
  lapply(s, function(lm){
    lapply(lm, function(y){
      lapply(y, function(x){
        car::Anova(x)['F value']
      })
    })
  })
})
npFval <- lapply(npFval, function(x){
  lapply(x, `[`, names(x[[1]]))
  apply(do.call(rbind, x), 2, as.list)
})
npFval <- lapply(npFval, function(s){
  lapply(s, function(x){
    lapply(x, `[`, names(x[[1]]))
    apply(do.call(rbind, x), 2, as.list)
  })
})
npFval <- lapply(npFval, function(s){
  lapply(s, function(ab){
    lapply(ab, function(t){
      do.call('cbind', t)
    })
  })
})
ppval <- mapply(function(pFvals, npFvals){
  mapply(function(x,y){
    mapply(function(r,s){
      mapply(function(p,q){
        rank(c(p,q))[1]/(n+1)
      }, p = as.list(data.frame(t(r))), q = as.list(data.frame(t(s))))
    }, r = x, s = y)
  }, x = pFvals, y = npFvals, SIMPLIFY = F)
}, pFvals = pFval, npFvals = npFval, SIMPLIFY = F)
ppval <- lapply(ppval, function(s){
  lapply(s, function(y){
    lapply(y, function(x){
      data.frame(t(x))
    })
  })
})
ppval <- lapply(ppval, function(s){
  lapply(s, function(y){
    data.frame(data.table::rbindlist(y, fill = T))
  })
})
ppval <- lapply(ppval, function(s){
  lapply(s, function(y){
    row.names(y) <- names(CWM.lm$C$abundance)
    return(y)
  })
})
var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
ppval <- lapply(ppval, function(s){
  lapply(s, function(ab){
    ab[,setdiff(var, names(ab))] <- NA
    ab <- ab[var]
    return(ab)
  })
})
ppval <- lapply(ppval, function(s){
  lapply(s, function(x){
    1 - x
  })
})
rm(pFval, npFval)

# Result tables
reg <- paste('~', paste(names(R$C), collapse = ' + '))
res <- mapply(function(CWM.lms, pvals, ppvals){
  mapply(function(CWM.lmab, pvalab, ppvalab){
    regression_table(CWM.lmab, reg, pvalab, ppvalab)
  }, CWM.lmab = CWM.lms, pvalab = as.list(data.frame(pvals)), ppvalab = ppvals, SIMPLIFY = F)
}, CWM.lms = CWM.lm, pvals = pval, ppvals = ppval, SIMPLIFY = F)
res <- lapply(res, function(s){
  mapply(function(ab, name){
    ab <- cbind(c(name, rep(' ', length(ab[,1])-1)), ab)
    names(ab)[1] <- 'Type'
    return(ab)
  }, ab = s, name = as.list(c('AB', 'PA')), SIMPLIFY = F)
})
res <- lapply(res, function(s){
  do.call('rbind', s)
})
res <- mapply(function(s, name){
  s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
  names(s)[1] <- 'Strata'
  return(s)
}, s = res, name = as.list(c('C', 'U')), SIMPLIFY = F)
res <- do.call('rbind', res)
kable(bold(res), row.names = F)
rm(CWM.lm, nCWM.lm, pval, ppval, reg, res) ; invisible(gc())
```

Traits convergence/divergence was measured with community weighted variances (CWV). Overall, observed CWVs did not differ from null expectations for all traits. We could found significant relationships between standardized effect size (SES) of CWVs and environmental variables (see table 4). Convergence of LES and LDMC and divergence of LA were found under south-west orientation. LA and LDMC values also diverged in communities with increasing curvature. Observed CVNNDs did not reveal niche differentiation among communities for all traits. But we also found significant relationships of their standardized effect size (SES) of CVNNDs with environmental variables (see table 4). niche differentiation of LA increased in lower curvature and wetness contexts, while differentiation of WD decreased in southwest orientation.

@@

```{r CWV}
# Observed CWVs
CWV <- mapply(function(Qs, CWMs, Ls){
  lapply(names(CWMs$abundance), function(v){
    apply(cbind(data.frame(CWMs$abundance[[which(names(CWMs$abundance) == v)]]),Ls$abundance), 1, function(x){
      sqrt(weighted.mean((Qs[,v]-x[1])^2, x[-1], na.rm=T))
    })
  })
}, Qs = Q, CWMs = CWM, Ls = L, SIMPLIFY = F)
CWV <- lapply(CWV, function(s){
  names(s) <- names(CWM$C$abundance)
  return(s)
})

# Simulated CWVs 
nL <- list(C = rep(list(L$C$abundance), n),
           U = rep(list(L$U$abundance), n))
nL <- lapply(nL, function(s){
  lapply(s,function(l){
    apply(l, 1, function(x){
      x[x != 0] <- sample(x[x != 0])
      return(x)
    })
  }) 
})
nL <- lapply(nL, function(s){
  lapply(s, function(x){
    t(x)
  })
})
nL <- lapply(nL, function(s){
  lapply(s, data.frame)
})
cl <- makeCluster(cores)
clusterExport(cl, list('nL', 'CWM', 'Q'))
nCWM.cwv <- clusterMap(cl, function(nLs, CWMs, Qs){
  lapply(nLs, function(nl){
    lapply(as.list(names(CWMs$abundance)), function(v){
      apply(nl, 1, function(x){
        weighted.mean(Qs[,v], x, na.rm=T)
      })
    })
  })
}, nLs = nL, CWMs = CWM, Qs = Q, SIMPLIFY = F)
stopCluster(cl)
nCWM.cwv <- lapply(nCWM.cwv, function(s){
  lapply(s, function(x){
    names(x) <- names(CWM$C$abundance)
    return(x)
  })
})
nCWV <- mapply(function(nLs, nCWM.cwvs, Qs){
  mapply(function(l,cwm){
  lapply(names(CWM$C$abundance), function(v){
    apply(cbind(data.frame(cwm[[which(names(cwm) == v)]]),l), 1, function(x){
      sqrt(weighted.mean((Qs[,v]-x[1])^2, x[-1], na.rm=T))
    })
  })
}, l = nLs, cwm = nCWM.cwvs, SIMPLIFY = F)
}, nLs = nL, nCWM.cwvs = nCWM.cwv, Qs = Q, SIMPLIFY = F)
nCWV <- lapply(nCWV, function(s){
  lapply(s, function(x){
    names(x) <- names(CWM$C$abundance)
    return(x)
  })
})
rm(nCWM.cwv, cl, nL)
invisible(gc())
## Non null CWVs
nCWV <- lapply(nCWV, function(s){
  lapply(s, `[`, names(s[[1]]))
  apply(do.call(rbind, s), 2, as.list)
})
nCWV <- lapply(nCWV, function(s){
  lapply(s, function(x){
    do.call('cbind', x)
  })
})
nCWV <- lapply(nCWV, function(s){
  lapply(s, data.frame)
})
CWV.pval <- mapply(function(CWVs, nCWVs){
  data.frame(
  mapply(function(x,y){
    mapply(function(p,q){
      rank(c(p,q))[1]/(n+1)
    }, p = as.list(x), q = as.list(data.frame(t(y))))
  }, x = CWVs, y = nCWVs)
)
}, CWVs = CWV, nCWVs = nCWV, SIMPLIFY = F)
CWV.pval.res <- lapply(CWV.pval, function(s){
  data.frame(
    inf = apply(apply(s, 2, p.adjust, method = 'fdr'), 2, function(x){
      length(which(x < 0.025))/length(x)
    }), 
    sup = apply(apply(1 - s, 2, p.adjust, method = 'fdr'), 2, function(x){
      length(which(x < 0.025))/length(x)
    }))
})
## CWV SES
CWV.ses <- mapply(function(CWVs, nCWVs){data.frame(
  mapply(function(x,y){
    mapply(function(p,q){
      (p - mean(q))/sd(q)
    }, p = as.list(x), q = as.list(data.frame(t(y))))
  }, x = CWVs, y = nCWVs)
)}, CWVs = CWV, nCWVs = nCWV, SIMPLIFY = F)
CWV.ses.lm <- mapply(function(s, Rs){
  apply(s, 2, function(y){
    lm(y ~., Rs)
  })
}, s = CWV.ses, Rs = R, SIMPLIFY = F)
CWV.ses.lm <- mapply(function(s, Rs){
  lapply(s, function(x){
    step(x, direction = 'both', trace = 0)
  })
}, s= CWV.ses.lm, Rs = R, SIMPLIFY = F)
CWV.ses.res <- lapply(CWV.ses.lm, function(s){
  lapply(s, function(x){
    summary(x)$coefficients
  })
})
CWV.ses.res <- lapply(CWV.ses.res, function(s){
  lapply(s, data.frame)
})
CWV.ses.res <- lapply(CWV.ses.res, function(s){
  lapply(s, function(x){
    x$res <- paste(format(x$Estimate, scientific = T, digits = 2, trim = T), apply(x[4], 1, stars, ns = '   '))
    return(x)
  })
})
CWV.ses.res <- lapply(CWV.ses.res, function(s){
  data.frame(data.table::rbindlist(lapply(s, function(x){
    data.frame(t(x[5]))
  }), fill = T))
})
var <- c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')
CWV.ses.res <- lapply(CWV.ses.res, function(s){
  s[,setdiff(var, names(s))] <- NA
  s <- s[var]
  s <- apply(s, 2, as.character)
  row.names(s) <- names(CWV.ses$C)
  s[is.na(s)] <- ''
  s <- data.frame(s)
  return(s)
})
CWV.ses.res <- mapply(function(s, CWV.ses.lms){
  s$`$R^2$` <- round(unlist(lapply(CWV.ses.lms, function(x){
    summary(x)$r.squared
  })),3)
  return(s)
}, s = CWV.ses.res, CWV.ses.lms = CWV.ses.lm, SIMPLIFY = F)
CWV.ses.res <- lapply(CWV.ses.res, function(s){
  s <- cbind(row.names(s), s)
  names(s)[1] <- 'Trait'
  return(s)
})
CWV.ses.res <- lapply(CWV.ses.res, function(s){
  sel <- which(unlist(lapply(as.list(data.frame(t(s))), function(t){
  any(grepl('*', t, fixed = T))
})))
  return(s[sel,])
})
CWV.ses.res <- mapply(function(s, name){
  s <- cbind(c(name, rep(' ', length(s[,1])-1)), s)
  names(s)[1] <- 'Strata'
  return(s)
}, s = CWV.ses.res, name = as.list(c('C', 'U')), SIMPLIFY = F)
CWV.ses.res <- do.call('rbind', CWV.ses.res)
rm(CWV.pval, CWV.pval.res, CWV.ses.lm, nCWV)
```

```{r CVNND}
# Observed CVNND
CVNND <- mapply(function(Ls, Qs){
  apply(Ls$abundance, 1, function(l){
  apply(Qs, 2, function(q){
    cati::CVNND(rep(q, l), na.rm = T)
    })
  })
}, Ls = L, Qs = Q, SIMPLIFY = F)

# Simulated CVNNDs 
cl <- makeCluster(cores)
clusterExport(cl, list('nQ', 'L'))
nCVNND <- clusterMap(cl, function(nQs, Ls){ 
  lapply(nQs, function(x){
    apply(Ls$abundance, 1, function(l){
      apply(x, 2, function(q){
        cati::CVNND(rep(q,l), na.rm = T)
      })
    })
  })
}, nQs = nQ, Ls = L)
stopCluster(cl)
rm(cl, nQ)
invisible(gc)

## Non null CVNNDs
nCVNND <- lapply(nCVNND, function(s){
  lapply(s, function(x){
    as.list(data.frame(t(x)))
  })
})
nCVNND <- lapply(nCVNND, function(s){
  lapply(s, function(x){
    lapply(x, function(y){
      names(y) <- colnames(CVNND$C)
      return(y)
    })
  })
})
nCVNND <- lapply(nCVNND, function(s){
  lapply(s, `[`, names(s[[1]]))
  apply(do.call(rbind, s), 2, as.list)
})
nCVNND <- lapply(nCVNND, function(s){
  lapply(s, function(x){
    do.call('cbind', x)
  })
})
nCVNND <- lapply(nCVNND, function(s){
  lapply(s, data.frame)
})
CVNND.pval <- mapply(function(CVNNDs, nCVNNDs){
  data.frame(
    mapply(function(x,y){
      mapply(function(p,q){
        rank(c(p,q))[1]/(n+1)
      }, p = as.list(x), q = as.list(data.frame(t(y))))
    }, x = as.list(data.frame(t(CVNNDs))), y = nCVNNDs)
  )
}, CVNNDs = CVNND, nCVNNDs = nCVNND, SIMPLIFY = F)
CVNND.pval <- lapply(CVNND.pval, function(s){
  row.names(s) <- colnames(CVNND)
  apply(s, 2, p.adjust, method = 'fdr')
  return(s)
})
CVNND.pval.res <- lapply(CVNND.pval, function(s){data.frame(
  inf = apply(s, 2, function(x){
    length(which(x < 0.05))/length(x)
  })
)})
CVNND.pval.res <- do.call('cbind', CVNND.pval.res)
names(CVNND.pval.res) <- names(CVNND.pval)
print(round(CVNND.pval.res, 3))
